---
title: "Assignment 2"
author: "Chadwick Reed"
date: "9/20/2020"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
    message: FALSE
    results: 'hide'
    warning: FALSE
    tidy: TRUE
    error: TRUE
---

```{r}
library(sf)
library(tidyverse)
library(ggthemes)
library(ggspatial)
library(units)
library(nngeo)
library(maps)
library(viridis)
library(RColorBrewer)
```

```{r}
townships <- st_read("https://opendata.arcgis.com/datasets/2f6e63652cbf4fdbb3499228b581db5d_27.kml", 
                  quiet = TRUE)

school_parcels <- st_read("https://opendata.arcgis.com/datasets/762b23b33fec4c57aea4356b58b21ebd_12.kml", 
                  quiet = TRUE) 

schools <- st_read("https://opendata.arcgis.com/datasets/7e57d5dc095442608ba84f67ef2cb7ad_1.kml", 
                  quiet = TRUE) 

college_parcels <- st_read("https://opendata.arcgis.com/datasets/7e829ed50de3423ca985a21a9de99d6f_12.kml", 
                  quiet = TRUE) 

colleges <- st_read("https://opendata.arcgis.com/datasets/ff6f0e1376a64d2e9a554efbaa737e95_11.kml", 
                  quiet = TRUE) 

parks <- st_read("https://opendata.arcgis.com/datasets/1da83b1a245c4d7884710502c1caf004_10.kml", 
                  quiet = TRUE)

crashes <- st_read("https://opendata.arcgis.com/datasets/2d42a003011a45718fc174446c7da76a_88.kml", 
                  quiet = TRUE)

libs <- st_read("https://opendata.arcgis.com/datasets/2ce0a345616a42cfba5bac2fb82d8980_16.kml", 
                  quiet = TRUE)
```

## Data by Type


__Polygon Data__

*A. Townships<br>
B.School parcels<br>
  B2. College parcels<br>
*


__Point Data__

*C. Parks<br>
D. Fatal or injurious crashes<br>
E. Schools<br>
  E2. Colleges<br>
  F. Libraries*


```{r}
NC_state_plane <- "+proj=lcc +lat_1=34.33333333333334 +lat_2=36.16666666666666 +lat_0=33.75 +lon_0=-79 +x_0=609601.22 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs  "

#I used the North Carolina State Plane Coordinate System (NAD1983, meters) for this assignment. 

townships <- townships %>%
  st_transform(NC_state_plane)

school_parcels <- school_parcels %>%
  st_transform(NC_state_plane)

schools <- schools %>%
  st_transform(NC_state_plane)

college_parcels <- college_parcels %>%
  st_transform(NC_state_plane)

colleges <- colleges %>%
  st_transform(NC_state_plane)

parks <- parks %>%
  st_transform(NC_state_plane)

crashes <- crashes %>%
  st_transform(NC_state_plane)

libs <- libs %>%
  st_transform(NC_state_plane)
```

All variables mapped below: 
```{r}
ggplot(townships) +
  geom_sf(alpha = .2) +
  geom_sf(data = school_parcels, fill = "Yellow3", color = "Yellow3") +
  geom_sf(data = college_parcels, fill = "Purple", color = "Purple") +
  geom_sf(data = crashes, color = "black", alpha = .1) +
  geom_sf(data = libs, color = "orange") +
  geom_sf(data = parks, color = "springgreen", size = .01) +
  theme_map()
```

## Points within a Polygon

### Figure 1: Schools within 100 meters of a fatal or injurious crash

```{r}
crashes_buffer <- st_buffer(crashes, dist = 100) %>%
  st_union()

ggplot(crashes_buffer) +
  geom_sf(shape = 8) +
  theme_map()
```

```{r}
school_crashes <- schools[crashes_buffer,]

ggplot(townships) +
  geom_sf(alpha = .2, color = NA, fill = "lightgray") +
  geom_sf(data=crashes_buffer, fill = "darkgray", color = "darkgray") +
  geom_sf(data=school_crashes,
          color="red",shape=17, size=2) +
  labs(caption = "Charlotte-Mecklenburg area schools within 100 meters of a fatal or injurious car crash") +
  theme_map()
```

```{r}
schools <- schools %>%
  st_join(school_crashes) %>%
  mutate(by_crash = !is.na(Name.y))


n_crashes_schools <- sum(schools$by_crash)

n_crashes_schools
```
There are 5 schools within 100 meters of a fatal or injurious crash site.

```{r}
n_schools<- length(schools$by_crash)

pct_crashes_schools <- n_crashes_schools / n_schools

pct_crashes_schools
```
About 2% of all schools are within 100 meters of a fatal or injurious crash site.


### Figure 2: Fatal or injurious car crashes within 100 meters of a school

```{r}
sp_buffer <- st_buffer(school_parcels, dist = 100, dissolve = FALSE, doEdge = TRUE) %>%
  st_union()

ggplot(sp_buffer) +
  geom_sf(fill = "yellow2", alpha = .25) +
  theme_map()
```

```{r}
crash_schools <- crashes[sp_buffer,]

ggplot(townships) +
  geom_sf(alpha = .2, color = NA, fill = "lightgray") +
  geom_sf(data=sp_buffer, fill = "yellow2", alpha = .75, color = "darkgray") +
  geom_sf(data=crash_schools,
          color="red",shape=17, size=2, alpha=0.45) +
  labs(caption = "Fatal or Injurious car crashes within 100m of schools in the Charlotte-Mecklenburg metro") +
  theme_map()
```

```{r}
crashes <- crashes %>%
  st_join(crash_schools) %>%
  mutate(by_school = !is.na(Name.y))


n_school_crashes <- sum(crashes$by_school)

n_school_crashes
```
There are 55 recorded fatal or injurious car crashes within 100 meters of Charlotte-Mecklenburg schools.


## Average Distance

### Figure 3: Proximity of parks to libraries 

```{r}
parks <- parks %>%
  mutate(libs_dist = st_nn(parks, libs, 
                                      returnDist = TRUE)$dist) %>%
            mutate(libs_dist = as.numeric(libs_dist))
```

```{r}
avg_libs_dist <- mean(parks$libs_dist)

avg_libs_dist
```
The average distance between libraries and parks is approximately 3091 meters.

```{r}
right_side <- st_bbox(parks)$xmax
left_side  <- st_bbox(parks)$xmin
top_side <- st_bbox(parks)$ymax
bottom_side <- st_bbox(parks)$ymin


ggplot(townships) +
  geom_sf() +
  geom_sf(data = parks,
          aes(color = avg_libs_dist)) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = TRUE) +
  scale_color_viridis_c(name = 
                          "Charlotte parks by distance to a library") +
  annotation_scale(location = "br")
```


## Overlapping Polygons

### Figure 4: Townships by presence of colleges

```{r}
townships <- townships %>%
  mutate(num_college_parcels = lengths(st_overlaps(townships, college_parcels))) %>%
  mutate(has_college_parcels = num_college_parcels > 0)

n_college_parcels_townships <- sum(townships$has_college_parcels)

#This returns 140. 140 townships within the Charlotte metro area have colleges.

left_side  <- st_bbox(college_parcels)$xmin
top_side <- st_bbox(college_parcels)$ymax

ggplot(college_parcels) +
  geom_sf(fill = "purple", color = NA) +
  geom_sf(data = townships,
          aes(fill = has_college_parcels)) +
  scale_fill_manual(values = c("7", "4"),
          name = "CLT Townships\nby presence of a college", 
          labels = c("Township without\na college",
                     "Township with a\ncollege")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side, 
           y = top_side - 1000, 
           label = paste(n_college_parcels_townships ,
                         "of the", 
                         length(townships$Name),
                         "Charlotte-area \ntownships neither contain\n nor overlap with", 
                         "a park."),
           hjust = .68, vjust = -.01, size = 2) +
  theme_map() +
  theme(panel.background = element_rect(fill = alpha("lightgray",0.5)),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```

## Number of Points

### Figure 5: Townships by number of fatal or injurious car crashes

```{r}
townships <- townships %>%
  mutate(num_crashes = lengths(st_covers(townships, crashes)))

ggplot(townships) +
  geom_sf(color = NA, 
          aes(fill = num_crashes)) +
  scale_fill_viridis_c( option = "viridis",
                        name = "CLT Townships\nby number of fatal or \ninjurious car crashes",
                       breaks = breaks <- seq(0, 1000, by = 20),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "crashes")) +
  annotation_scale(location = "bl") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(legend.position = c(0.6,0), legend.background = element_rect(fill = alpha("white", .5), color = "gray"))
```

## Density

### Townships by educational density

```{r}
townships <- townships %>%
  mutate(area = set_units(st_area(townships), km^2)) %>%
  mutate(num_schools = lengths(st_covers(townships, schools))) %>%
  mutate(schools_dens = as.numeric(num_schools / area))

ggplot(townships) +
  geom_sf(color = NA, 
          aes(fill = num_schools)) +
  
    scale_fill_gradientn(colors = brewer.pal( 5, "YlGn"),
                       breaks = breaks <- seq(0, 1, by = .05),
                       labels = paste(prettyNum(breaks, big.mark = ","))) +
  labs(caption = "Charlotte-Mecklenburg townships by educational density") +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tl",
                         style = north_arrow_minimal())
```

