---
title: "Assignment 7"
author: "Chadwick Reed"
date: "11/2/2020"
output:
  html_document:
    toc: true
    toc_depth: 1
    toc_float: true
    theme: "paper"
---

I grew up in Charlotte, where the closest major university was UNC Charlotte, which was about 30 minutes from my parents' home. So, when I moved to Minneapolis, one of the first things I noticed was the size of the student population in the city. In addition to the University of Minnesota (UMN is one of the country's largest universities), a number of fairly prominent private colleges exist in the area. However, in terms of academic presence, the Twin Cities still pale in comparison to Boston-Cambridge. Having lived in Cambridge for a few months now, I have a decent idea of the higher education landscape here, so I thought it would be interesting to look colleges and universities on the other side of the river.

```{r load libraries, message=FALSE, results='hide', echo=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(leaflet)
library(htmltools)
library(htmlwidgets)
library(raster)
library(gstat)
library(spatial)
library(dplyr)
```

```{r, message=FALSE, results='hide', echo=FALSE}
c_plane <- "+proj=longlat +datum=WGS84"

MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"

hoods <- st_read("https://opendata.arcgis.com/datasets/3525b0ee6e6b427f9aab5d0a1d0a1a28_0.kml", quiet = TRUE) %>%
  dplyr::select(Name) %>%
  st_transform(c_plane)

higher_ed <- read_csv("C:/Users/chad/Documents/R/creed1-vis/[vis]assignment7_files/higher_ed.csv") 

charge_st <- read_csv("C:/Users/chad/Documents/R/creed1-vis/[vis]assignment7_files/charge_st.csv")
```


# Preliminary Map

```{r, message=FALSE, error=FALSE, echo=FALSE, warning=FALSE}
leaflet(hoods) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(highlightOptions = highlightOptions(fillColor = "gray", 
                                                  fillOpacity = .5),
              label = ~Name, 
              weight = 1) %>%
  addCircleMarkers(data = higher_ed, color = "Yellow", radius = 4, opacity = .9, label = higher_ed$Name)
```


# Chloropleth

```{r, message=FALSE, error=FALSE, echo=FALSE, warning=FALSE}
higher_ed_sf <- st_as_sf(higher_ed, coords=c("Longitude", "Latitude")) %>%
  st_set_crs(4326) %>%
  st_transform(c_plane)

hoods_higher_ed <- hoods %>%
  mutate(num_higher_ed = lengths(st_contains(hoods, higher_ed_sf)))
```
```{r, message=FALSE, error=FALSE, echo=FALSE, warning=FALSE}
hoods_higher_ed$label <-
  paste(hoods_higher_ed$Name, "<br>",
        hoods_higher_ed$num_higher_ed, "colleges or universities") %>%
  lapply(htmltools::HTML)

bins <- seq(min(hoods_higher_ed$num_higher_ed),
            max(hoods_higher_ed$num_higher_ed), by=5)
            
pal <- colorNumeric("inferno", 
                    domain = hoods_higher_ed$num_higher_ed,
                    na.color = "#00000000")

leaflet(hoods_higher_ed) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(highlightOptions = highlightOptions(fillOpacity = 0.5),
              label = ~label,
              fillColor = ~pal(num_higher_ed),
              weight = 1, color = "black") %>%
  addLegend(pal = pal,
            values = ~num_higher_ed,
            bins = 3,
            opacity = 0.75,
            title = "Colleges/Universities per Neighborhood",
            position = "topright")

```


# Points

```{r, message=FALSE, error=FALSE, echo=FALSE, warning=FALSE}
hood_points <- st_centroid(st_transform(hoods_higher_ed, crs = MA_state_plane)) %>%
  st_transform(c_plane)

leaflet(hood_points) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addCircles(label = ~label, fillColor = ~pal(num_higher_ed), stroke = FALSE, radius = 200, fillOpacity = 0.5) %>%
  addControl("Number of colleges/universities",
             position = "topright")
```


# Interpolation

```{r, message=FALSE, error=FALSE, echo=FALSE, warning=FALSE, results='hide'}
hood_points_sp <- hood_points %>%
  st_transform(MA_state_plane) %>%
  as_Spatial()

hoods_poly_sp <- hoods %>%
  st_transform(MA_state_plane) %>%
  as_Spatial

hoods_raster <- raster(hoods_poly_sp, res=10)

gs <- gstat(formula = num_higher_ed~1, locations = hood_points_sp)
idw_interp <- interpolate(hoods_raster, gs)
```
```{r, message=FALSE, error=FALSE, echo=FALSE, warning=FALSE}
idw_interp_clip <- mask(idw_interp, hoods_poly_sp)

leaflet(hood_points) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addRasterImage(idw_interp_clip, colors = pal, opacity = 0.75) %>%
  addLegend(pal = pal,
            values = ~num_higher_ed,
            bins = 3,
            opacity = 0.75,
            title = "Number of colleges/universities",
            position = "topright")
```


# Discussion

Of the maps above, I believe that the final - the interpolation map - is the most effective at clearly conveying the spatial distribution of colleges and universities in Boston. I think it's also the most visually interesting. It's certainly the most straightforward; one needs only to read the title to immediately understand the data.  As such, I think it's also the most appropriate. The least compelling of the three iterations is definitely the point/centroid map. While the chloropleth is superior to the point map in terms of legibility and appropriateness, I think it is weakened by the extent to which it can effectively convey intensity. With the heatmap, you get a good idea of the magnitude of the difference between, say, the Fenway neighborhood, with its 15 institutions, and South Boston, which has zero. Perhaps this is an issue with the color palette, but while the chloropleth communicates that the Fenway neighborhood has the greatest concentration, the degree of intensity of the variations between neighborhoods is lost.