---
title: "Assignment 7"
author: "Chadwick Reed"
date: "11/2/2020"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: "paper"
---

## Loading Libraries

```{r load libraries, message=FALSE, results='hide', echo=FALSE}
library(tidyverse)
library(sf)
library(leaflet)
library(htmltools)
library(htmlwidgets)
library(raster)
library(gstat)
library(spatial)
library(dplyr)
```

## Designating Dataframes and Projections

I grew up in Charlotte, where the closest major university was UNC Charlotte, which was about 30 minutes from my parents' home. So, when I moved to Minneapolis, one of the first things I noticed was the size of the student population in the city. In addition to the University of Minnesota (UMN is one of the country's largest universities), a number of fairly prominent private colleges exist in the area. However, in terms of academic presence, the Twin Cities still pale in comparison to Boston-Cambridge. Having lived in Cambridge for a few months now, I have a decent idea of the higher education landscape here, so I thought it would be interesting to look colleges and universities on the other side of the river.

```{r, message=FALSE, results='hide', echo=FALSE}
c_plane <- "+proj=longlat +datum=WGS84"

MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"

hoods <- st_read("https://opendata.arcgis.com/datasets/3525b0ee6e6b427f9aab5d0a1d0a1a28_0.kml", quiet = TRUE) %>%
  dplyr::select(Name) %>%
  st_transform(c_plane)

higher_ed <- read_csv("C:/Users/chad/Documents/R/creed1-vis/[vis]assignment7_files/higher_ed.csv") 

charge_st <- read_csv("C:/Users/chad/Documents/R/creed1-vis/[vis]assignment7_files/charge_st.csv")
```


## Maps

### Preliminary

```{r}
leaflet(hoods) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(highlightOptions = highlightOptions(fillColor = "gray", 
                                                  fillOpacity = .5),
              label = ~Name, 
              weight = 1) %>%
  addCircleMarkers(data = higher_ed, color = "Yellow", radius = 4, opacity = .9, label = higher_ed$Name)
```


### Chloropleth

```{r}
higher_ed_sf <- st_as_sf(higher_ed, coords=c("Longitude", "Latitude")) %>%
  st_set_crs(4326) %>%
  st_transform(c_plane)

hoods_higher_ed <- hoods %>%
  mutate(num_higher_ed = lengths(st_contains(hoods, higher_ed_sf)))
```
```{r}
hoods_higher_ed$label <-
  paste(hoods_higher_ed$Name, "<br>",
        hoods_higher_ed$num_higher_ed, "colleges or universities") %>%
  lapply(htmltools::HTML)

bins <- seq(min(hoods_higher_ed$num_higher_ed),
            max(hoods_higher_ed$num_higher_ed), by=1)
            
pal <- colorNumeric("inferno", 
                    domain = hoods_higher_ed$num_higher_ed,
                    na.color = "#00000000")

leaflet(hoods_higher_ed) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(highlightOptions = highlightOptions(fillOpacity = 1),
              label = ~label,
              fillColor = ~pal(num_higher_ed),
              weight = 1, color = "black") %>%
  addLegend(pal = pal,
            values = ~num_higher_ed,
            bins = 3,
            opacity = 0.75,
            title = "Colleges/Universities per Neighborhood",
            position = "topright")

```


## Points

```{r}
hood_points <- st_centroid(st_transform(hoods_higher_ed, crs = MA_state_plane)) %>%
  st_transform(c_plane)

leaflet(hood_points) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addCircles(label = ~label, fillColor = ~pal(num_higher_ed), stroke = FALSE, radius = 200, fillOpacity = 0.5) %>%
  addControl("Number of colleges/universities",
             position = "topright")
```


### Interpolation

```{r, message=FALSE, results='hide', echo=FALSE}
hood_points_sp <- hood_points %>%
  st_transform(MA_state_plane) %>%
  as_Spatial()

hoods_poly_sp <- hoods %>%
  st_transform(MA_state_plane) %>%
  as_Spatial

hoods_raster <- raster(hoods_poly_sp, res=10)

gs <- gstat(formula = num_higher_ed~1, locations = hood_points_sp)
idw_interp <- interpolate(hoods_raster, gs)
```
```{r}
idw_interp_clip <- mask(idw_interp, hoods_poly_sp)

leaflet(hood_points) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addRasterImage(idw_interp_clip, colors = pal, opacity = 0.75) %>%
  addLegend(pal = pal,
            values = ~num_higher_ed,
            bins = 3,
            opacity = 0.75,
            title = "Number of colleges/universities (approximated)",
            position = "topright")
```


## Discussion

Of the three representations above I believe that the data-as-points representation is the most appropriate portrayal of the data and the raster interpolation is the least appropriate, despite it being a visually interesting map. I chose to represent percentage data per neighborhood; this is already an average that does not take into account variations within neighborhoods, which could be significant. Portraying the data through centroids is a level of abstraction which forces the viewer to recognize it is aggregate data fo the neighborhood. On the other hand, showing this education data as a continuous surface is misleading - these values are discontinuous.

General trends in the maps show that the closer a neighborhood is to the Cambridge/Downtown Boston/Charles River area of the map, the higher the percentage of the neighborhood’s residents that have bachelor’s degrees or higher. A raster interpolation surface represents this to some extent, but is deceptive because it projects “expected values” that may be very far from reality.
