---
title: "Assignment 4"
author: "Chadwick Reed"
date: "10/7/2020"
output: html_document
---

## Introduction

In this exercise, I looked at parks and recreational areas in the city of Wake Forest, North Carolina. Initially, I was going to look at Charlotte, NC, which is where I spent most of my childhood. However, in terms of physical size/square mileage, Charlotte is massive, and my analyses were taking way too long. I chose Wake Forest because I went to summer camp at Wake Forest University for several years, and the data was easily accessible.

```{r}
library(dplyr)
library(knitr)
library(osmdata)
library(opentripplanner)
library(tidyverse)
library(sf)
library(ggthemes)
library(ggspatial)
library(rosm)
```

```{r}
WF_rec <- st_read("https://opendata.arcgis.com/datasets/03ff718a97ab4bb998403160581993f8_0.kml")
```


## Open Trip Planner Setup

```{r}
opq(bbox = 'Wake Forest NC USA') %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_xml(file = 'OTP/graphs/default/wf_streets.osm')
```

```{r}
NC_state_plane <- "+proj=lcc +lat_1=34.33333333333334 +lat_2=36.16666666666666 +lat_0=33.75 +lon_0=-79 +x_0=609601.22 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"

wf_street_features <- opq(bbox = 'Wake Forest NC USA') %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_sf()

wf_streets <- wf_street_features$osm_lines %>%
  st_transform(crs = NC_state_plane)
```


### Map of Wake Forest

```{r}
ggplot(wf_streets) +
  geom_sf() +
  theme_map()
```

```{r}
path_otp <- otp_dl_jar("OTP")
```

```{r}
path_data <- file.path(getwd(), "OTP")
path_otp <- paste(path_data, "otp.jar",sep = "/")

otp_build_graph(otp = path_otp, dir = path_data, memory =1024)
```

```{r}
otp_setup(otp = path_otp, dir = path_data, memory =1024)
```

```{r}
otpcon <- otp_connect()
```


## Isochrones

```{r}
iso_10min_walk <- 
  otp_isochrone(otpcon = otpcon, fromPlace = WF_rec, 
                mode = "WALK", cutoffSec = 600) %>%
  st_transform(crs = NC_state_plane) %>%
  mutate(mode = "walk")

iso_10min_drive <- 
  otp_isochrone(otpcon = otpcon, fromPlace = WF_rec, 
                mode = "CAR", cutoffSec = 600) %>%
  st_transform(crs = NC_state_plane) %>%
  mutate(mode = "drive")

iso_all_modes <- rbind(iso_10min_drive, iso_10min_walk)

otp_stop()
```


```{r}
right_side <- st_bbox(iso_all_modes)$xmax
left_side  <- st_bbox(iso_all_modes)$xmin
top_side <- st_bbox(iso_all_modes)$ymax
bottom_side <- st_bbox(iso_all_modes)$ymin

ggplot(iso_all_modes) +
  annotation_map_tile(zoomin = 0, progress = "none") +
  geom_sf(aes(fill = mode), alpha = 0.5) +
  geom_sf(data = WF_rec) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_fill_viridis_d(option = "plasma", name = "Area that is reachable within 5 minutes",
                       labels = c("By car", "By foot")) +
  theme_map() +
  labs(caption = "Basemap Copyright OpenStreetMap contributors")
```


## Area Calculations

```{r}
iso_areas <- iso_all_modes %>%
  mutate(area = st_area(iso_all_modes)) %>%
  st_set_geometry(NULL) %>%
  pivot_wider(names_from = mode, values_from = area) 

ggplot(iso_areas, 
       aes(x = as.numeric(walk), y = as.numeric(drive))) +
  geom_point() +
  scale_x_continuous(name = 
            "Area within a five-minute walking distance\nof a park or recreation space\n(square km)",
            breaks = breaks <- seq(10000, 130000, by = 20000),
            labels = breaks / 1000000) +
  scale_y_continuous(name = 
            "Area within a five-minute driving distance\nof a park or recreation space\n(square km)",
            breaks = breaks <- seq(0, 700000, by = 100000),
            labels = breaks / 1000000) +
  theme_bw()
```

